Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-24 paper/reviewer_response_PNAS.tex

Of note, position 1 for the H1 HA in the supplementary figure is not even shown.

\response{
The reviewer is correct. 
Figure~S3 shows data for a H1 HA as measured previously by Doud and Bloom (2016), and these prior data are simply presented here in H3 numbering for comparative purposes (the original Doud and Bloom paper used a different HA numbering scheme that makes site-to-site comparisons with the current study difficult). 
The Doud and Bloom study used mutant libraries in which all codons in HA \emph{except} for the one encoding the N-terminal methionine were mutagenized.
Therefore, that study did not estimate the effects of mutations to the first codon, which is why it is not shown in Figure~S3.
We have added a sentence to the legend for this supplementary figure explaining this fact.
}

2) While other important features of the HA like several cysteines seem to be conserved, there is no pressure on other elements which are considered important for viral fitness including the NGT glycosylation site in the beginning of HA1 which is conserved in all group 2 HAs or the adjacent NAT glycosylation site which is conserved in H3, H7, H10 and H15. On the other hand, head glycosylation sites, which are highly variable in nature, seem to be conserved. Again, viruses without these glycosylation sites should be rescued and evaluated for fitness relative to wild type.

\response{This is a valid point, and we have now also individually evaluated the reverse-genetics titers of three glycosylation site mutants.
Two of the mutants, T24F and T40V, remove the aforementioned highly conserved but mutationally tolerant NGT and NAT glycosylation sites near the start of HA1.
The third mutation, S287A, removes a glycosylation site that our experiments suggest is under functional constraint.
The new Figure S5 shows that the T24F and T40V mutants grow to titers comparable to wildtype, supporting the conclusion that they are expendable at least in cell culture. 
However, the titer of the S287A mutant is modestly lower than wildtype, supporting the conclusion that this glycan contributes to viral growth in cell culture.

One possible explanation is simply that cell culture does not fully capture the constraints on HA in nature.
But a more interesting possibility is that conservation in nature is not expected to strictly parallel mutational tolerance.
The tendency for a site to change in nature depends on how much selection there is for the mutations.
For instance, Wei et al (\textit{Science Translational Medicine}, 2:24ra21, 2010) have pointed out that glycosylation motifs evolved rapidly in the globular head of HA from human H1 influenza in the period from 1918 to 2008 (seasonal H1N1), but changed very little in the same period in the swine H1 lineages descended from the 1918 virus.
The likely reason is differential strengths of selection: human influenza is under strong pressure from accumulated immune memory, whereas swine influenza is under less immune pressure because pigs are short lived and so are often naive when infected with influenza.
This fact highlights that even mutationally tolerant sites will only change rapidly if there is a benefit to doing so, while even somewhat constrained sites can change if there is sufficient pressure.
A strength of our experiments is that we have controlled the selection pressure (selection only for viral growth), and so can measure mutational tolerance in the absence of the confounding effects of immunity.

In addition to the data in Figure S5, we have added the following text to the Results:

\begin{quote}
\textsl{
The disparity between the relative conservation of these sites in nature and their mutational tolerance in our study could be because cell culture does not fully capture the constraints on HA function in nature, or could be because these sites are not under strong immune pressure and so mutations at them are not positively selected in nature.}
\end{quote}
}

@@

\response{The reviewer has a number of questions and suggestions about the experimental design.
Some of these suggestions are things that were already implemented in our experiments and we simply failed to explain in adequate detail, while others are things that we have evidence are unimportant. 
Below we describe the issues one-by-one.

In addition, we have now included our \emph{full original lab notes} for the relevant experiments (\url{https://github.com/jbloomlab/Perth2009-DMS-Manuscript/tree/master/lab_notes}) so that readers can examine everything that we did in full detail.
This level of transparency goes well beyond that usually taken in scientific papers, and will enable readers to fully understand our reasoning and procedures for each detail of the experiments.
}

a) The TCID50 is likely underestimating the number of virus particles with HA genomic segments because not all of them (likely the minority) will be capable of inducting a full infectious cycle (but their presence might still be detected in the cell supernatants).

\response{The reviewer is certainly correct that TCID50 assays underestimate the number of HA genomic segments, as it is well known that the number of physical influenza virions exceeds the number of infectious units by 10-100 fold.
This is why we used an extremely low MOI (0.0035) so that most cells are still infected by no more than one virion.
It is impossible to grow a viral stock that is completely free of defective particles.
However, in both our original rescue with helper virus and in the passage, we follow the steps suggested by Carolina Lopez and Scott Hensley (PMID 27047455) to minimize defective particles.
These include low MOI (0.0035 TCID50 per cell is quite low) and short growth times to avoid accumulation of defective particles (the helper-virus rescue was allowed to propagate for just 24 hours, and the low-MOI passage for just 48 hours).

@@

In short, the presence of non-infectious particles is an inherent feature of all virology work.
The best that can be done is to use MOIs and growth times that ensure that they do not dominate the experiments, and we have done that.
}

b) The infectious inoculum is never washed off the infected cells meaning that any particles that are transferred from the rescue transfection are going to be detected during sequencing. This might even apply to naked RNPs released from dead cells.

\response{
We apologize for not expanding on this detail in the Methods of the original paper.

In fact, like the reviewer, we initially thought that a media change to wash away the inoculum might be important. 
Therefore, in our experiments, we performed such a media change to remove the initial inoculum for biological replicates 1 and 2, and for one of the technical replicates (3-2) of biological replicate 3.
We did not perform such a change for the other technical replicate (3-1) of biological replicate 3.
It turns out that the media change makes no discernible difference, which is why we neglected to expand on this point in the original manuscript. 
We now have added the following explanatory text to the Methods:

@@

\response{
We apologize if the original wording of our methods were unclear. 
For each mutant virus library, we passaged a total of $9\times10^5$ TCID50, but this was passaged across \emph{fifteen} 15-cm dishes \emph{each} of which received 25 ml.
Because we used an inoculum of 2.5 TCID50 / $\mu$l across these fifteen dishes, this corresponds to dilutions that range from 1:37 to 1:294.
The number of dishes used to passage each mutant virus library is now mentioned in the Methods; before it had to be calculated from the stated total number of cells and the cells per dish.

Also, we note that the supernatants from the helper virus rescues are collected at just 24 hours post-infection, which leads to lower viral titers (which is why we have to passage relatively large volumes), but also presumably much lower amounts of DIs / RNA.
This is probably the reason why (as described above) removing the inoculum did not discernibly affect the results.
}

d) The supernatant is never centrifuged before RNA extraction. Cell debris is not removed which is likely contributing significantly to RNA sequences detected in the supernatant. However, RNA associated with cell debris is not representative of RNA from fit, replicating viruses.

\response{We apologize for failing to provide sufficiently detailed methods stating that we ultra-centrifuged \emph{clarified} the viral supernatant before RNA extraction.
The supernatant was clarified by centrifugation at 2000 $\times$ \textit{g} for five minutes to remove any cellular debris.
We have updated the Methods to now include this detail, and (as mentioned above) also now include our full lab notes so that all other details can be reviewed.
Specifically, see \url{https://github.com/jbloomlab/Perth2009-DMS-Manuscript/blob/master/lab_notes/Perth09_HA_Library_Rescue.pdf} (relevant details of the notebook can be found in pages 21-25).
}

@@

The experiments should be repeated with at least one more passaging step, proper removal of inoculum and washing of the cells after infection. Centrifugation steps should also be included between the passages and before RNA extraction in the end.

\response{For the reasons described above, it is not necessary to repeat the experiments.
We do not deny that there are slightly different ways that the experiments could be done, but in the end, the ``proof is in the pudding.'' 
In the manuscript, we show that the measurements made using the experimental parameters that we chose lead to useful data as judged by the following criterion:
\begin{itemize}
\item The results from fully independent experimental replicates are highly correlated (Fig.~1D), which would not be the case if the results were dominated by noise.
\item The experimental measurements can inform substitution models that describe influenza virus's evolution much better than standard models (Table~1).
\item The experimental measurements provide information that can help predict the fates of influenza virus lineages in nature (Fig.~5).
\end{itemize}
For any experimental study there are an almost infinite number of slight methodological variations that could be employed, and it is impossible to test them all.
Our experiments used reasonable parameter choices, were performed in full biological triplicate, and give meaningful / useful results with respect to the larger goals of the study.}

@@

\response{We individually validated viral growth of a C52A mutant, which removes a Cys critical in forming a disulfide bond with Cys-277 and maintaining HA structure.
We confirmed that this mutant grows very poorly when rescued, and mutations at site 52 never appear in sequenced isolates.

As another example, mutation N170H fails to reach $>$1\% frequency in nature, and according to our dataset, is a mutation towards a highly unpreferred amino acid. 
}

3) The authors picked 31 random clones to evaluate the mutation rate per clone. These sequences should be made available. Did this set include viruses without a proper start codon (see above)? What was their growth phenotype in vitro?

\response{We have now added the list of mutations from the 31 clones to Dataset S3.
Two of the clones had a mutation that removed the ATG start codon.
We did not test the growth of these clones. 
As explained in Figure~S2, these were plasmid variants chosen randomly to be Sanger sequenced to evaluate the mutation rate prior to creating virus libraries; they were not evaluated as viruses.
The new Figures S4 to S6 now show the growth phenotype of other selected mutants, one of which includes a start codon mutant.
}

@@

\subsubsection*{Major points:}

The authors themselves acknowledge the most important limitation of this study: that the experimental measure of viral growth in cell culture is not reflective of true fitness in nature. Most importantly, the work presented here does not capture host-induced pressures on the antigenic regions of the viral protein. This is evidenced by the discrepancy between the high mutational tolerance they calculated for the stalk region and the actual change in stalk amino acid sequence. A simple protein BLAST comparison between the amino acid sequences of HA from H3N2 viruses Perth09 and HK68 shows a 77\% conservation in head amino acids but a 94\% conservation in stalk amino acids. Additionally, according to figure 2, there is no observable reversion of the G78D mutation they introduced despite the dominance of G78 in the population as seen in figure S1. The authors recognize this in their work and do not attempt to exaggerate the significance of these findings. Still, it would be useful to differentiate between head and stalk mutations when depicting individual amino acid mutations in figure 4 and their frequency trajectories to highlight that the predominance of mutations that occur in nature are in the head.

\response{We thank the reviewer for acknowledging our efforts to fairly present the considerable caveats associated with cell-culture measurements.

The suggestion to make a version of Fig.~4 that distinguished between head and stalk mutations is a good one.
It is difficult to also overlay the structural location (stalk versus head) on the current Fig.~4, since colors are already used to indicate mutational effects.
We have therefore created a new supplemental figure (Fig.~S11) that shows these data---as the reviewer suspected, this figure shows that head mutations reach high frequencies much more commonly than stalk mutations.

@@

The reviewer is correct that G78D appears to be a site where the correspondence between selection in our experiments and frequency in nature is low.
This makes sense, as G78D is one of the two sites that we specifically identified as being a lab adaptation (see Figure~S1).
We now explicitly mention this fact in the Discussion as follows:

\begin{quote}
\textsl{Indeed, a vast amount of work in virology has chronicled the ways in which experiments can select for lab artifacts or fail to capture important pressures that are relevant in nature.
As an example, although we identified G78D as favorable for viral growth in cell culture, this mutation never fixes in nature. }
\end{quote}
}

The data regarding greater mutational tolerance in the stalk compared to the head should be paired with data from deep mutational scanning utilizing stalk- and head-specific antibodies as a selection process as this group has previously done with WSN H1 (Doud et al., Nat Comm 2018, Doud et al., PLoS Path 2017) to see if tolerant sites are responsible for viral escape.

\response{We agree that these would be very interesting experiments to perform.
However, such experiments would be far beyond the scope of our study.
As stated in the Abstract and Significance Statement, the main goal of our study is to perform deep mutational scanning of a recent H3 HA and to use the data to better predict H3N2 viral evolution in nature.
A large set of additional experiments comparing stalk- and head-specific antibodies in H3 and H1 HAs is clearly beyond the scope of this goal.
}

@@

The correlation for prospective analysis (post-Perth) is smaller than the correlation for post-hoc analysis (pre-Perth). Since prospective analysis is more important when predicting evolutionary fates, it might be beneficial for the authors to comment on this difference.

\response{
The reviewer is correct that on the complete data sets in the original submission, the correlation was slightly higher for the post-hoc (post-Perth) and prospective (pre-Perth) analysis.
But at the suggestion of Reviewer \#1, we have now also included a post-Perth analysis using \emph{only} HA sequences from unpassaged viral isolates (there are not enough unpassaged isolates from before 2009 to perform a similar unpassaged pre-Perth analysis).
As shown in the new Figure 5, the correlation for the unpassaged post-Perth is actually higher than either other set, demonstrating that there is not a systematic trend for poorer performance in prospective analyses.
}

A number of issues interfere with the clarity of this manuscript, particularly when appealing to a broader scientific audience. While this is understandable given the nature of this study, several adjustments should be made. A clear description of mutational effect is never provided in the results, though it forms a crucial component of the data. As such, it is not really intuitive as to why the range from -10-0 should be considered "neutral." Additionally, mutational tolerance is quantified as Shannon entropy of the re-scaled amino-acid preferences at each site, but it is unclear how this entropy is calculated. The entire first half of the second results section describing how the experimentally informed codon substitution model they generate is better than existing models and the associated Table 1 is very difficult to understand. AIC is never explained. The significance of dN/dS being much less than 1 for conventional models should be explained. The purpose of modeling of ExpCM, site avg. is unclear.

\response{We appreciate these suggestions from the reviewer and have made the following changes:
\begin{itemize}
\item We have now included a description of how we calculated mutational effects in the Results section.
\item We agree that a cutoff of -10 to 0 is arbitrary and removed the sentence ``Mutations that reach high frequencies generally have neutral or beneficial effects according to our experimental measurements." 
\item We included in the Methods a description of how we calculated Shannon entropy.
\item We now define AIC in the legend of Table 1.
\item We have now included an explanation of dN/dS $\ll$1 for conventional substitution models.
\item We now describe how the ``ExpCM, site avg.'' substitution model is a control that shows that the improvements of the ExpCM are due to the site-specific nature of the data, since the ``ExpCM, site avg.'' model averages data across sites.
\end{itemize}
}

### 2018-06-24 paper/reviewer_response_PNAS.tex

\newcommand{\comment}[1]{{\color{red}[\textsl{#1}]}}
\newcommand{\response}[1]{{\color{black}#1}}

\title{Response to reviews of ``Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants'' for \textit{PNAS}}
\author{Juhye M. Lee, John Huddleston, Michael B. Doud, Kathryn A. Hooper,\\Nicholas C. Wu, Trevor Bedford, Jesse D. Bloom}

\begin{document}
\maketitle

We thank the reviewers for the careful reading of our manuscript.
We were gratified that they recognized the quality of our work: one of the reviewers said, ``the experimental setup is sound and there is strong correlation among technical and biological replicates.''

The main contribution of our work is to use show that deep mutational scanning of an H3 HA provides information that is informative about the evolutionary fates of H3N2 influenza virus variants in nature.
We appreciated the reviewers' recognition of the important step our work takes in linking experiments with quantitative models of influenza virus evolution, and their recognition that we carefully discuss the caveats as well as the strengths of our approach.
Reviewer \#1 did make an important suggestion to look at HA sequences only from unpassaged virus isolates---in the revised version we have done that, and found that our data are even more predictive of evolutionary fate for these higher quality sequences.

Most of the reviewer questions had to do with more tangential aspects of our work, such as validation of specific mutations and phylogenetic analyses.
In the revised version, we have addressed all of these questions.
Therefore, we think that you will find that the revisions have further strengthened the paper.

@@

\response{We are sorry that the reviewer found some parts of the manuscript confusing.
We have revised the parts of the text that the reviewer highlighted, performed additional validation experiments, and clarified the experimental setup.
These modifications should fully address the reviewer's concerns.}

\subsubsection*{Major points}

1) It is disturbing that there is no preference for having a methionine in the first position of the HA ORF. The authors discuss this and refer to one 2002 paper that shows additional start sites for non-human H2. This is a pretty weak explanation. If this is really the case the authors should actually isolate/rescue these viruses and test their fitness relative to wild type.

\response{
We have now used reverse genetics to generate viruses carrying a mutation of the initiating methionine to lysine.
These start-codon mutant viruses grow to appreciable titers, suggesting that there is indeed some flexibility in the site of translation initiation in this HA.
These data are in Figure S4 of the revised manuscript.
As we note in the manuscript, this finding is concordant with prior work on suggesting additional translation initiation sites sites in H2 HAs.
Because translation initiation of HA is not the focus of our manuscript, we have not investigated the exact mechanism or identity of the alternate start sites.
However, the data we have added clearly show that the canonical ATG is not strictly necessary for viral growth.
}

@@

Of note, position 1 for the H1 HA in the supplementary figure is not even shown.

\response{
The reviewer is correct. 
Figure~S3 shows data for a H1 HA as measured previously by Doud and Bloom (2016), and these prior data are simply presented here in H3 numbering for comparative purposes (the original Doud and Bloom paper used a different HA numbering scheme that makes site-to-site comparisons with the current study difficult). 
The Doud and Bloom study used mutant libraries in which all codons in HA \emph{except} for the one encoding the N-terminal methionine were mutagenized.
Therefore, that study did not estimate the effects of mutations to the first codon, which is why it is not shown in Figure~S3.
We have added a sentence to the legend for this supplementary figure explaining this fact.
}

2) While other important features of the HA like several cysteines seem to be conserved, there is no pressure on other elements which are considered important for viral fitness including the NGT glycosylation site in the beginning of HA1 which is conserved in all group 2 HAs or the adjacent NAT glycosylation site which is conserved in H3, H7, H10 and H15. On the other hand, head glycosylation sites, which are highly variable in nature, seem to be conserved. Again, viruses without these glycosylation sites should be rescued and evaluated for fitness relative to wild type.

\response{We have now also individually evaluated the reverse-genetics titers of three glycosylation site mutants.
Two of the mutants, T24F and T40V, remove the aforementioned highly conserved but mutationally tolerant NGT and NAT glycosylation sites near the start of HA1.
The third mutation, S287A, removes a glycosylation site that our experiments suggest is under functional constraint.
The new Figure S5 shows that the T24F and T40V mutants grow to titers comparable to wildtype, supporting the conclusion that they are expendable at least in cell culture. 
However, the titer of the S287A mutant is modestly lower than wildtype, supporting the conclusion that this glycan contributes to viral growth in cell culture.

One possible explanation is simply that cell culture does not fully capture the constraints on HA in nature.
But a more interesting possibility is that conservation in nature does not strictly parallel mutational tolerance.
The tendency for a site to change in nature depends on how much selection there is for the mutations.
For instance, Wei et al (\textit{Science Translational Medicine}, 2:24ra21, 2010) have pointed out that glycosylation motifs evolved rapidly in the globular head of HA from human H1 influenza in the period from 1918 to 2008 (seasonal H1N1), but changed very little in the same period in the swine H1 lineages descended from the 1918 virus.
The likely reason is differential strengths of selection: human influenza is under strong pressure from accumulated immune memory, whereas swine influenza is under less immune pressure because pigs are short lived and so are often naive when infected with influenza.
This fact highlights that even mutationally tolerant sites will only change rapidly if there is a benefit to doing so, while even somewhat constrained sites can change if there is sufficient pressure.
A strength of our experiments is that we have controlled the selection pressure (selection only for viral growth), and so can measure mutational tolerance in the absence of the confounding effects of immunity.

@@

We are excited about the future directions that the reviewer suggests.
However, the current study is a complete and important contribution even if (as is almost invariably the case in science) future work can probably take things further.
Researchers have been performing functional experiments on influenza virus for decades, and have been trying to forecast the virus's evolution for decades---yet ours is the first study to successfully use such functional experiments to predict the success of viral strains in nature. 
We have gone to great pains to fairly discuss limitations and caveats and how things could be improved in the future, but that should not detract from the advances in the current paper.
}

\subsubsection*{Minor Points:}

The correlation for prospective analysis (post-Perth) is smaller than the correlation for post-hoc analysis (pre-Perth). Since prospective analysis is more important when predicting evolutionary fates, it might be beneficial for the authors to comment on this difference.

\response{
The reviewer is correct that on the complete data sets in the original submission, the correlation was slightly higher for the post-hoc (post-Perth) and prospective (pre-Perth) analysis.
But at the suggestion of Reviewer \#1, we have now also included a post-Perth analysis using \emph{only} HA sequences from unpassaged viral isolates (there are not enough unpassaged isolates from before 2009 to perform a similar unpassaged pre-Perth analysis).
As shown in the new Figure 5, the correlation for the unpassaged post-Perth is actually higher than either other set, suggesting that there may not a systematic trend for poorer performance in prospective analyses.
}

### 2018-06-24 paper/manuscript.tex
filler spots for additional codon mutations

\subsection*{Creation of MDCK-SIAT1-TMPRSS2 cells}
When growing influenza virus in cell culture, trypsin is normally added to cleave HA into its mature form.
To obviate the need for trypsin, we engineered MDCK-SIAT1 cells and MDCK-SIAT1-CMV-PB1 cells to constitutively express the TMPRSS2 protease, which cleaves and activates HA in the human airways~\cite{bottcher2006proteolytic,bottcher2010cleavage}.
The human TMPRSS2 cDNA ORF was ordered from OriGene (NM\_005656) and cloned into a pHAGE2 lentiviral vector under an EF1$\alpha$-Int promoter followed by an IRES driving expression of mCherry to create plasmid pHAGE2-EF1aInt-TMPRSS2-IRES-mCherry-W.
We used the lentiviral vector to transduce MDCK-SIAT1 or MDCK-SIAT1-CMV-PB1 cells, and sorted an intermediate mCherry-positive population by flow cytometry.
We refer to the sorted bulk population as MDCK-SIAT1-TMPRSS2 cells or MDCK-SIAT1-CMV-PB1-TMPRSS2 cells.
There is no selectable marker for the TMPRSS2; however, we maintain the cells at low passage number, and have seen no indication that they lose their ability to support the growth of viruses with H3 HAs in the absence of exogenous trypsin.

@@

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.4\textwidth]{figs/supp_point_mut_validation/startcodon_titers.pdf}}
\caption{
{\bf Validation that the Perth/2009 HA is somewhat tolerant to mutation of the canonical start codon.}
We created variants of the Perth/2009 HA in which the canonical start codon was mutated (amino-acid mutation M(-16)K, codon mutation ATG(-16)AAA), a conserved cysteine was mutated to alanine (C52A, codon mutation TGC52GCA), and the same cysteine was synonymously mutated (C52C, codon mutation TGC52TGT).
We selected these mutations because our deep mutational scanning results in Figure~\ref{fig:logoplot} surprisingly suggest that the start codon (position -16) is fairly tolerant of mutations, but that site 52 is highly intolerant of mutation to anything other than cysteine.
The synonymous C52C mutation is a negative control mutation that is not expected to have any effect.
We then used reverse genetics to generate viruses carrying the wildtype HA or each point mutant, with GFP packaged in the PB1 segment to enable easy titering by flow cytometry~\cite{bloom2010permissive,hooper2013mutant}.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
As expected, the C52A virus is essentially non-viable, while the C52C mutation is roughly equivalent to wildtype.
The M(-16)K mutation is only moderately attenuated, validating that the canonical start codon is not completely essential for viral growth.
}
\label{suppfig:startcodon_validation}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.4\textwidth]{figs/supp_point_mut_validation/glycan_titers.pdf}}
\caption{\label{suppfig:glycan_validation}
{\bf Validation of viral variants with mutations at N-linked glycosylation motifs.}
We created variants of the Perth/2009 HA with mutations that eliminated N-linked glycosylation motifs (Asn-Xaa-Ser/Thr) at asparagine residues 22, 38, and 285 (these are mutations T24F, T40V, and S287A, respectively).
We then used reverse genetics to generate viruses carrying the wildtype HA or each of these mutants.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The viruses with mutations at the motifs at residues 22 and 38 reached titers at least as high as wildtype, whereas the virus with a mutation to the motif at residue 285 was modestly attenuated.
}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.45\textwidth]{figs/supp_point_mut_validation/tm_titers.pdf}}
\caption{\label{suppfig:TM_validation}
{\bf Validation of the mutational tolerance of a site in the transmembrane domain.}
We created a variant of the Perth/2009 HA with a transmembrane domain mutation, C199(HA2)K, at a site that our deep mutational scanning suggests should be highly mutationally tolerant (Figure~\ref{fig:logoplot}).
We then used reverse genetics to generate viruses carrying the wildtype HA or this mutant.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The virus with the mutation reached titers comparable to wildtype.
}
\end{suppfigure}

>>>>

\subsection*{Creation of MDCK-SIAT1-TMPRSS2 cells}
When growing influenza virus in cell culture, trypsin is normally added to cleave HA into its mature form.
To obviate the need for trypsin, we engineered MDCK-SIAT1 cells and MDCK-SIAT1-CMV-PB1~\cite{bloom2010permissive} cells to constitutively express the TMPRSS2 protease, which cleaves and activates HA in the human airways~\cite{bottcher2006proteolytic,bottcher2010cleavage}.
The human TMPRSS2 cDNA ORF was ordered from OriGene (NM\_005656) and cloned into a pHAGE2 lentiviral vector under an EF1$\alpha$-Int promoter followed by an IRES driving expression of mCherry to create plasmid pHAGE2-EF1aInt-TMPRSS2-IRES-mCherry-W.
We used the lentiviral vector to transduce MDCK-SIAT1 or MDCK-SIAT1-CMV-PB1 cells, and sorted an intermediate mCherry-positive population by flow cytometry.
We refer to the sorted bulk population as MDCK-SIAT1-TMPRSS2 cells or MDCK-SIAT1-CMV-PB1-TMPRSS2 cells.
There is no selectable marker for the TMPRSS2; however, we maintain the cells at low passage number, and have seen no indication that they lose their ability to support the growth of viruses with H3 HAs in the absence of exogenous trypsin.

@@

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.4\textwidth]{figs/supp_point_mut_validation/startcodon_titers.pdf}}
\caption{
{\bf Validation that the Perth/2009 HA is somewhat tolerant to mutation of the canonical start codon.}
We created variants of the Perth/2009 HA in which the canonical start codon was mutated (amino-acid mutation M(-16)K, codon mutation ATG$\rightarrow$AAA), a conserved cysteine was mutated to alanine (C52A, codon mutation TGC$\rightarrow$GCA), and the same cysteine was synonymously mutated (C52C, codon mutation TGC$\rightarrow$TGT).
We selected these mutations because our deep mutational scanning results in Figure~\ref{fig:logoplot} surprisingly suggest that the start codon (position -16) is fairly tolerant of mutations, but that site 52 is highly intolerant of mutation to anything other than cysteine.
The synonymous C52C mutation is a negative control mutation that is not expected to have any effect.
We then used reverse genetics to generate viruses carrying the wildtype HA or each point mutant, with GFP packaged in the PB1 segment to enable easy titering by flow cytometry~\cite{bloom2010permissive,hooper2013mutant}.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
As expected, the C52A virus is essentially non-viable, while the C52C mutation is roughly equivalent to wildtype.
The M(-16)K mutation is only moderately attenuated, validating that the canonical start codon is not completely essential for viral growth.
}
\label{suppfig:startcodon_validation}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.4\textwidth]{figs/supp_point_mut_validation/glycan_titers.pdf}}
\caption{\label{suppfig:glycan_validation}
{\bf Validation of viral variants with mutations at N-linked glycosylation motifs.}
We created variants of the Perth/2009 HA with mutations that eliminated N-linked glycosylation motifs (Asn-Xaa-Ser/Thr) at asparagine residues 22, 38, and 285 (these are mutations T24F, T40V, and S287A, respectively).
\comment{The codon mutations were NNN$\rightarrow$NNN, NNN$\rightarrow$NNN, and NNN$\rightarrow$NNN, respectively.}
We then used reverse genetics to generate viruses carrying the wildtype HA or each of these mutants.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The viruses with mutations at the motifs at residues 22 and 38 reached titers at least as high as wildtype, whereas the virus with a mutation to the motif at residue 285 was modestly attenuated.
}
\end{suppfigure}

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.45\textwidth]{figs/supp_point_mut_validation/tm_titers.pdf}}
\caption{\label{suppfig:TM_validation}
{\bf Validation of the mutational tolerance of a site in the transmembrane domain.}
We created a variant of the Perth/2009 HA with a transmembrane domain mutation, C199(HA2)K (\comment{codon mutation NNN$\rightarrow$NNN}), at a site that our deep mutational scanning suggests should be highly mutationally tolerant (Figure~\ref{fig:logoplot}).
We then used reverse genetics to generate viruses carrying the wildtype HA or this mutant.
Each variant was generated in duplicate, and the plots show the viral titer in the supernatant at 53 hours post-transfection.
The virus with the mutation reached titers comparable to wildtype.
}
\end{suppfigure}
